A Global Picture of Food Insecurity and Hunger

### 2020-09-15 docs/modeling_fies.tex
added modeling in tex

\subsection{Covariates}
We model subnational rates of severe and moderate-to-severe food insecurity as a function of several covariates related to human development, food security, health, infrastructure, and the environment (See Table \ref{tab:covars}.  We aimed for covariates that had subnational spatial resolution or peer-reviewed projections available through the year 2030.  For many variables, we combined historic subnational data with national projections through the 2020s to derive simple subnational estimates for the next decade.  For projections based on the Shared Socioeconomic Pathways (SSP) framework, we used the middle-of-the-road pathway, SSP2, and for climatological variables, we use forecasts based on Representative Concentration Pathway 6.  Finally, we adjusted several variables to account for the effects of COVID-19, including GDP per capita using June 2020 estimates from the World Bank \citep{prospects2020}, stunting and wasting using August 2020 estimates published in \textit{The Lancet} \citep{headey2020impacts}, as well as the poverty headcount index (CITATION).  For a detailed overview of the steps involved in processing and preparing each covariate, see the Appendix.

@@

Random Forest

Hyperparameter selection

@@

\section{Modeling Covariates Into the Future}
\subsection{Annualized Rate of Change}
For many of our covariates, to extrapolate recent trends to the future, it is necessary to draw on annualized rates of change, which are then used to anticipate the rate of future change.  This method can be used for any variable that is a fraction or a rate.  We use this method to estimate future rates of stunting, wasting, and malaria mortality, as well as to model future admin shares of GDP and population.  Thus, we give an overview of the methodology here.

@@

>>>>

\subsection{Covariates}
We model subnational rates of severe and moderate-to-severe food insecurity as a function of several covariates related to human development, food security, health, infrastructure, and the environment (See Table \ref{tab:covars}).  We aimed for covariates that had subnational spatial resolution or peer-reviewed projections available through the year 2030.  For many variables, we combined historic subnational data with national projections through the 2020s to derive simple subnational estimates for the next decade.  For projections based on the Shared Socioeconomic Pathways (SSP) framework, we used the middle-of-the-road pathway, SSP2, and for climatological variables, we use forecasts based on Representative Concentration Pathway 6.  Finally, we adjusted several variables to account for the effects of COVID-19, including GDP per capita using June 2020 estimates from the World Bank \citep{prospects2020}, stunting and wasting using August 2020 estimates published in \textit{The Lancet} \citep{headey2020impacts}, as well as the poverty headcount index (CITATION).  For a detailed overview of the steps involved in processing and preparing each covariate, see the Appendix.

@@

Given the subnational FIES data resulting from the disaggregation combined with the covariates stated above (See Table \ref{tab:covars}), we modelled subnational rates of severe and moderate-to-severe food insecurity for the whole world in the period of 2010 to 2030. We conduced suitable statistical methods such as Least Absolute Shrinkage and Selection Operator (LASSO) regression and Bayesian Stochastic Search and Variable Selection (Bayes SSVS), as well as machine learning methods such as Random Forest regression. Based on various model validation metrics, Random Forest (RF) regression performs best which can be seen as a contribution to the increasing importance of such methods in the field (CITATION: Master Thesis M\"uller?).

By using the most popular model validation metrics, we observed a in-sample mean squared error (MSE) of 0.00?? with a R2 of 0.999? for the RF regression model. Since these models are prone to overfitting (CITATION) it is crucial to check the out-of-sample prediction performance. Hence, we trained the RF regression model with 75\% of the data and predicted rates of severe and moderate-to-severe food insecurity for the remaining 25\%. We got an out-of-sample MSE of 0.00?? with a R2 of 0.99??. This excellent out-of-sample performance makes the RF regression model particularly suitable for our application. See the Supplement Info for a detailed description and validation of all models.

In RF regression models two main hyperparameters can be selected. One parameter which describes the number of variables randomly selected as candidates for splitting a node and another which describes the forest average number of unique cases (data points) in a terminal node, “mtry” and “nodesize” respectively (CITATION, manual r package randomforestSRC). We calculated the Out-Of-Bag (OOB) error for different combinations of the hyperparameters. The OOB error is the mean prediction error of the observation not included in the bootstrap aggregating or bagging (CITATION). By choosing the combination of hyperparameters with the lowest OOB error we found the optimal values which we used in our RF regression model to predict subnational rates of severe and moderate-to-severe food insecurity.

@@

\section{Predicting Covariates Into the Future}
\subsection{Annualized Rate of Change}
For many of our covariates, to extrapolate recent trends to the future, it is necessary to draw on annualized rates of change, which are then used to anticipate the rate of future change.  This method can be used for any variable that is a fraction or a rate.  We use this method to estimate future rates of stunting, wasting, and malaria mortality, as well as to model future admin shares of GDP and population.  Thus, we give an overview of the methodology here.

@@

\section{Description and Validation of Models}

\begin{figure}[H]
  \centering
  \includegraphics[width=\linewidth]{img/model/residuals.png}
  \caption{In-Sample Residuals of the RF regression model. Panel (A) shows the Moderate-to-Severe model and Panel (B) shows the Severe model.}
\end{figure}
